PM Mungbean meta-analysis

### 2020-06-22 02_Preliminary_analysis.Rmd
Move AUDPC plotting to Preliminary analysis

>>>>

### AUDPC visualise raw data

First lets check we have all the AUDPC observations.

```{r AUDPC_check}
PM_MB_means %>%
   filter(is.na(AUDPC_m)) %>%
   distinct(trial_ref, location, year,AUDPC_m, first_sign_disease, final_assessment)%>%
   kable(align = "c")
```

The trials `r unique(PM_MB_means[is.na(PM_MB_means$AUDPC_m),"trial_ref"]) ` don't have an estimate of AUDPC and need to be filled in. 
The following script estimates AUDPC from the final disease severity rating and the first sign of disease information.

```{r SingleObsAUDPC}
source(here("R/AUDPC_for_single_obs.R"))
write.csv(PM_MB_means,
          "cache/1911_PM_MB_means&Ygains.csv", row.names = FALSE)
```

>How well does 'area under the disease progress curve' (AUDPC) predict grain yield?

and _Does spray_management scenario mitigate yield loss through reducing AUDPC?_

We would expect to see grain yield decline with increasing AUDPC. Let's just plot yield vs AUDPC.

```{r Yield_vs_AUDPC}

plot(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))

There appears to be a weak positive relationship between the two continuous variables.  
Does a linear model confirm this?

```{r AUDPClm}
summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))
```

This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

Lets tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship. 
Let's do the plot again for each trial where there are no missing values.  
Note: The plots are ordered by mean yield of the trial.

```{r AUDPCxTrial}
sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPC_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPC_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(PM_MB_means$grain_yield.t.ha., na.rm = TRUE)),
      xlim = c(0, max(PM_MB_means$AUDPC_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPC_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

}
```

This sequence of graphs show that AUDPC and disease incidence tends not impact on yields if the season is a low yielding, ie lower than 1 tonne per hectare.  
The higher the yield the greater greater the influence on the disease on the yield especially if the crop is high yielding with AUDPC also being large enough to reduce yield.

Let's have a look again at the same sequence of graphs and analyses for area under the disease progress _stairs_.

```{r Yield_vs_AUDPS}
plot(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

summary(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i,]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_means$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_means$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control", ],
          col = "red")

}

```
The plots from the AUPDS is similar to the AUDPC, however according to the linear model the AUDPC is a better, significant fit.

A mixed effect model would be more suitable to investigate the interactive effects of location, season, and trial. 
The mixed effect model will help us decide on the best approach and structure for our meta-analysis model.
